---
title: "David's work"
author: "David Skrill"
date: "4/17/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
fancy.summarize <- source("fancy_summarize.R")$value
```

```{r}
d <- readRDS("austin.rds")
d <- d %>% mutate(vehicle_year = ifelse(vehicle_year<1960,1960,vehicle_year)) %>% 
  mutate(vehicle_year = ifelse(vehicle_year > 2017,NA, vehicle_year)) %>% 
  mutate(vehicle_make = fct_lump_n(vehicle_make,n = 25)) %>% 
  mutate(vehicle_model = fct_lump_n(vehicle_model,n = 250))
```

```{r}
d.numeric.all <- d %>% select(subject_age,subject_sex,frisk_performed,
                          search_conducted, search_person,
                          search_vehicle)
d.numeric.search <- d %>% filter(search_conducted==T) %>% 
  select(contraband_found,contraband_drugs,contraband_weapons,frisk_performed)

d.cat <- d %>% select(subject_race,search_basis,reason_for_stop,
                      vehicle_make,vehicle_model,vehicle_registration_state,
                      vehicle_year)
```

```{r}
summary.stats1 <- d.numeric.all %>% mutate_all(as.numeric) %>% 
  mutate(subject_sex = subject_sex - 1) %>%  #defaults to 1/2 coding
  as.data.frame() %>% 
  fancy.summarize(.,nmis=T,uniq=T)
```

```{r}
summary.stats2 <- d.numeric.search %>% 
  mutate_all(as.numeric) %>% 
  as.data.frame() %>% 
  fancy.summarize(uniq=T,nmis=T)
```

```{r}
summary.stats.tables <- apply(d.cat,2,tabyl)
summary.stats.tables
```

# Investigating the Hit Rate
```{r}
d %>% filter(search_conducted==T) %>% tabyl(subject_race)
```

Because there are relatively few stops of Asian/Pacific Islanders or people of other races, I will combine these. Since we are primarilly interested in the differences in hit rates by race, I will remove the relatively few instances in which race is unknown.

```{r}
d2 <- d %>% filter(subject_race != "unknown") %>% 
  mutate(subject_race = fct_lump_min(subject_race,20000))
```

```{r}
d2 %>% 
  filter(search_conducted) %>% 
  group_by(subject_race) %>% 
  summarise(
    hit_rate = mean(contraband_found, na.rm = T)
  )

d2 %>% 
  filter(search_conducted) %>% 
  group_by(subject_race) %>% 
  summarise(
    hit_rate = mean(frisk_performed, na.rm = T)
  )
```

I will investigate the hit rate at the individual officer level. There are many officers with very few (or only one) stop. There is a very long tail, with one officer making over 10000 stops (averaging over 3 per day over the study period)! There is a similar pattern in the number of searches conducted. I will restrict the analysis to officers with 12 or more searches (corresponding to officers in or above the 75th percentile).

```{r}
stops.all <- d2 %>% tabyl(officer_id_hash)
hist(stops.all$n,breaks=200)
quantile(stops.all$n)
```

```{r}
searches.all <- d2 %>% filter(frisk_performed==T) %>% 
  tabyl(officer_id_hash)
hist(searches.all$n,breaks=200)
quantile(searches.all$n)

```


```{r}
quantile(searches.all$n,seq(0,1,.1))
ids.to.keep <- searches.all %>% filter(n>18) %>% pull(officer_id_hash)
searches <- d2 %>% filter(officer_id_hash %in% ids.to.keep) %>%
  filter(search_conducted==T) %>% 
  mutate(hit = (contraband_found | contraband_drugs | contraband_weapons)) %>% 
  select(officer_id_hash, contains("subject_"), 
         reason_for_stop,frisk_performed,search_conducted,hit,contains("contraband"))
head(searches)
```

```{r}
# hit_rates <- searches %>%
#   group_by(officer_id_hash,subject_race) %>%
#   summarise(
#     hit_rate = mean(contraband_found, na.rm = T)
#   )
hit_rates <- searches %>% 
  group_by(officer_id_hash,subject_race) %>% 
  summarise(
    hit_rate = mean(frisk_performed, na.rm = T)
  )


hit_rates <- hit_rates %>% 
  filter(subject_race %in% c("black", "white", "hispanic")) %>% 
  spread(subject_race, hit_rate, fill = 0) %>% 
  rename(white_hit_rate = white) %>% 
  gather(minority_race, minority_hit_rate, c(black, hispanic)) %>%
  arrange(officer_id_hash)
hit_rates
```

```{r}
# We'll use this just to make our axes' limits nice and even
max_hit_rate <- hit_rates %>% ungroup %>% 
  select(ends_with("hit_rate")) %>% 
  max()
hit_rates %>% 
  ggplot(aes(
    x = white_hit_rate,
    y = minority_hit_rate
  )) +
  geom_point() +
  # This sets a diagonal reference line (line of equal hit rates)
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  # These next few lines just make the axes pretty and even
  scale_x_continuous("White hit rate", 
    limits = c(0, max_hit_rate + 0.01),
    labels = scales::percent
  ) +
  scale_y_continuous("Minority hit rate", 
    limits = c(0, max_hit_rate + 0.01),
    labels = scales::percent
  ) +
  # This makes sure that 1% on the x-axis is the same as 1% on the y-axis
  coord_fixed() +
  # This allows us to compare black v. white and Hispanic v. white side by
  # side, in panels
  facet_grid(. ~ minority_race)
  # Depending on your version of ggplot2, you may be able to use the syntax 
  # below (the newer ggplot2 syntax)---which is clearer, in my opinion.
  # But older versions of ggplot2 will only accept the above syntax
  # facet_grid(cols = vars(minority_race))
```

```{r}
searches <- d2 %>% filter(officer_id_hash %in% ids.to.keep) %>%
  filter(subject_race %in% c("black","hispanic","white")) %>% 
  mutate(subject_race = relevel(subject_race,ref="white")) %>% 
  mutate(subject_race = fct_drop(subject_race)) %>%
  filter(frisk_performed==T) %>% 
  mutate(hit = (contraband_found | contraband_drugs | contraband_weapons)) %>% 
  select(officer_id_hash, contains("subject_"), 
         reason_for_stop,frisk_performed,search_conducted,hit,contains("contraband"))

hit_rates_binary <- searches %>%
  filter(subject_race %in% c("black","hispanic","white")) %>% 
  mutate(subject_race = fct_drop(subject_race)) %>% 
  mutate(subject_race = relevel(subject_race,ref="white")) %>% 
  group_by(officer_id_hash,subject_race) %>% 
  summarise(
    hit_rate = mean(contraband_found, na.rm = T),
    nsearches = n(),
    nhits = sum(contraband_found,na.rm=T)
  )
table(hit_rates_binary$subject_race)
```

```{r}
library(rstanarm)
summary_stats <- function(posterior) {
  x <- invlogit(posterior)  # log-odds -> probabilities
  t(apply(x, 2, quantile, probs = c(0.025, 0.5, 0.975)))
}
shift_draws <- function(draws) {
  sweep(draws[, -1], MARGIN = 1, STATS = draws[, 1], FUN = "+")
}


```

```{r,eval=F}
SEED <- 101
wi_prior <- normal(-1.3, 1) #the overall hit rate is .298; log(.298) ~ -1.21
stanfit1 <- stan_glmer(cbind(nhits, nsearches-nhits) ~ 1 + (1 +subject_race | officer_id_hash), 
                             data = hit_rates_binary, family = binomial("logit"),
             prior_intercept = wi_prior, seed = SEED)
```

```{r}
stanfit1 <- readRDS("stanfit1.rds")
alphas <- shift_draws(as.matrix(stanfit1))
post1 <- summary_stats(alphas)
post1 <- post1[-nrow(post1),]
```

```{r}
officer.medians <- post1[1:(nrow(post1)-5), 2]
median.mat <- matrix(officer.medians,nrow=length(unique(ids.to.keep)),
                          ncol=3,byrow = T)
mu <- apply(median.mat,1,mean)
output = rep(0,130)
for (ii in 1:130){
  temp = mean(median.mat[ii,])
  ssr = sum((median.mat[ii,] - mu[ii])^2)
  output[ii] = ssr
}
```


```{r}
# fit2 <- stan_glmer(cbind(nhits, nsearches-nhits) ~ (1 | officer_id_hash),
#                              data = hit_rates_binary, family = binomial("logit"),
#              prior_intercept = wi_prior, seed = SEED)
# alphas <- shift_draws(as.matrix(fit2))
# partialpool <- summary_stats(alphas)
# partialpool <- partialpool[-nrow(partialpool),]
# 
# fit3 <- stan_glmer(cbind(nhits, nsearches-nhits) ~ subject_race + (1 | officer_id_hash),
#                              data = hit_rates_binary, family = binomial("logit"),
#              prior_intercept = wi_prior, seed = SEED)
# alphas <- shift_draws(as.matrix(fit3))
# partialpool <- summary_stats(alphas)
# partialpool <- partialpool[-nrow(partialpool),]}
# SEED <- 101
# wi_prior <- normal(-1.2, 1) #the overall hit rate is .298; log(.298) ~ -1.21
# fit_partialpool <-stan_glmer(cbind(nhits, nsearches-nhits) ~ (subject_race | officer_id_hash),
#                              data = hit_rates_binary, family = binomial("logit"),
#              prior_intercept = wi_prior, seed = SEED)
```




```{r}
# logit1 <- glm(hit ~ subject_race, data = searches, family="binomial")
# summary(logit1)
# ```
# 
# It is likely the case that the hit rate will vary substantially by officer.
# ```{r}
# lmer1 <- lme4::glmer(hit ~ 1 + (1|subject_race),
#                      data = searches, family="binomial")
# summary(lmer1)
# ```
# 
# ```{r}
# lmer2 <- lme4::glmer(hit ~1 +  subject_race + (1|officer_id_hash),
#                      data = searches, family="binomial")
# summary(lmer2)
# ```
# 
# ```{r}
# lmer3 <- lme4::glmer(hit ~1 + (1 +subject_race | officer_id_hash),
#                      data = searches, family="binomial")
# summary(lmer3)
```
 


